---
title: Multiple Regression Analysis of Meningococcal Infection two Populations from
  1992-2015
author: "George Boateng-Sarfo and Dylan Miska"
date: "11/22/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---
##Library
```{r, message=FALSE}
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function
library("scatterplot3d") # Used to show 3d scatter plot
```

##Get the Data
```{r, message=FALSE}
# getwd()
# setwd()
Epid = read.csv("data1.csv", header = TRUE)
head(Epid)
```


##Make an Epid matix
```{r, message=FALSE}
E <- as.matrix(cbind(1, Epid$CP, Epid$APPSIMD))
head(E)
```

##Parameter Estimates (β̂)
```{r, message=FALSE}
y<- (Epid$IMD)
y<- as.numeric(y)
y
beta.hat <-solve(t(E) %*% E) %*% t(E) %*% y
beta.hat
```

##Predicted value
```{r, message=FALSE}
mu.hat <- E %*% beta.hat
head(mu.hat)
```
##Find Variance( σ2), Residual Standard Error, Degree of Freedom, Variance Covariance Matrix.
```{r, message=FALSE}
n<-nrow(E)
p<-ncol(E)

sigma2<-sum((y-mu.hat)^2/(n-p))
sigma2
```

##Residual Standard Error
```{r, message=FALSE}
sqrt(sigma2)
```

##Degree of Freedom
```{r, message=FALSE}
n-p
```

##Variance Covariance Matrix
```{r, message=FALSE}
Var<- solve(t(E) %*% E) *sigma2
Var
```

##Standard errors of the parameter estimates
```{r, message=FALSE}
sqrt(diag(Var))
```

##t-value for the parameter estimates
```{r, message=FALSE}
t.values<-beta.hat/sqrt(diag(Var))
t.values
```

##P-values for the parameter estimates
```{r, message=FALSE}
2*(1-pt(abs(t.values),n-p))
```
##Checking the result with lm
```{r, message=FALSE}
result.lm<-lm(IMD~CP+APPSIMD, data = Epid)
result.lm
```

```{r, message=FALSE}
summary(result.lm)
```

```{r, message=FALSE}
plot(result.lm)
```


```{r, message=FALSE}
anova(result.lm)
```
```{r, message=FALSE}
CP <- Epid$CP
APPSIMD <- Epid$APPSIMD
IMD <- Epid$IMD

s3d <- scatterplot3d(CP, APPSIMD, IMD, type = "h", color = "blue", angle=55, pch = 16)
s3d$plane3d(result.lm)
```


##Model for children (0-14 years)
```{r, message=FALSE}
Epid2<- read.csv("data2.csv", header = T)
Epid2 = Epid2 [-26:-25,]
head(Epid2)
```

##Matrix for Children Population
```{r, message=FALSE}
E2 <- as.matrix(cbind(1, Epid2$PC, Epid2$APPSIMD))
head(E2)
```

##Estimate for Children Population
```{r, message=FALSE}
y2<- (Epid2$MID)
head(y2)
beta.hat2 <- solve(t(E2) %*% E2) %*% t(E2) %*% y2
beta.hat2
```
##Predicted values for Children 
```{r, message=FALSE}
mu.hat2 <- E2 %*% beta.hat2
head(mu.hat2)
```

##Find covariance in the children Population
```{r, message=FALSE}
n2<-nrow(E2)
p2<-ncol(E2)

sigma2c<-sum((y-mu.hat2)^2/(n2-p2))
sigma2c
```

##Residual Standard Error for the children population
```{r, message=FALSE}
sqrt(sigma2c)
```

##Degree of fredom for the children Population
```{r, message=FALSE}
n-p
```

##Generating varaince covariance matrix for children Population 
```{r, message=FALSE}
Varc<- solve(t(E2) %*% E2) *sigma2c
Varc
```

##Standard Errors of the parameters Estimated for the children population
```{r, message=FALSE}
sqrt(diag(Varc))
```
##t-values for the estimated paramters fot the children 
```{r, message=FALSE}
t.valuesc<-beta.hat2/sqrt(diag(Varc))
t.valuesc
```
##P-values  for the estimated paramters for the children
```{r, message=FALSE}
2*(1-pt(abs(t.valuesc),n2-p2))
```

##Checking the results with the lm funx
```{r, message=FALSE}
result.lm_Children<-lm(MID~PC+APPSIMD, data = Epid2)
result.lm_Children
```
##Summary of the model for the children Population
```{r, message=FALSE}
summary(result.lm_Children)
```

```{r, message=FALSE}
plot(result.lm_Children)
```


```{r, message=FALSE}
CP <- Epid2$PC
APPSIMD <- Epid2$APPSIMD
IMD <- Epid2$MID

s3d <- scatterplot3d(CP, APPSIMD, IMD, type = "h", color = "blue", angle=55, pch = 16)
s3d$plane3d(result.lm_Children)
```
